---
title: "Data Import"
output: github_document
---
```{r setup, echo = FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```


This document will show how to import data.


# Import the FAS litters CSV

```{r}
litters_df = read_csv("data/FAS_litters.csv")

litters_df = janitor::clean_names(litters_df)
```

# Look at the dataset
```{r}
litters_df

head(litters_df )
tail(litters_df,10 )
```


```{r eval = FALSE}
view(litters_df)
```

# Learning assessment

Import `FAS_pups.csv` in.

```{r}
pups_df = read.csv("data/FAS_pups.csv")

pups_df = janitor::clean_names(pups_df)

pups_df

```


Use absolute path

```{r}
pups_df = read.csv("/Users/caiqi/Desktop/ds /data_wrangling/data_wrangling_1/data/FAS_pups.csv")
```


#Look at read_csv options

col_names and skipping  rows

```{r}
litters_df =
  read_csv(
    file = "data/FAS_litters.csv",
    skip = 2
  )
```

What about missing data

```{r}
litters_df =
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", ".",""),
  )
```


What if we code `group` as a factor variable?

```{r}
litters_df =
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", ".",""), 
    col_types = cols(
      Group = col_factor()
    )
  )
```


# Import Excel file

Import MLB 2011 summary data.

```{r}
mlb_df = read_excel("data/mlb11.xlsx", sheet = "mlb11")
```


# Import SAS data

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
```


# Never use read.csv()

```{r}
litters_df = read.csv("data/FAS_litters_df.csv")
```

Never do this either:

```{r}
litters_df$L
```




